Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-11 popl24/automation.tex
checkpoint

\section{Automating Security Verification}

>>>>

\section{Automating Security Verification}
\label{section-automation}

\subsection{Brute Force Computation of Program Distributions}
\label{section-bruteforce}

We first explore a brute force approach to automatically verifying
program behaviour, the first step of which is to compute the basic
distribution of a program $\prog$. The program distribution
$\progd(\prog)$ can obviously be automatically derived from that, as
can be any of its marginal or conditional distributions. In this
Section, we consider two automated mechanisms for deriving
$\progd(\prog)$ in this way. The first is through a straightforward
computation of all final program memories.  The second approach
rewrites $\minifed$ programs to stratified Datalog programs, that are
amenable HPC optimizations such as parallelization and GPU matrix
computations as shown in recent work
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

\begin{fpfig}[t]{Truth table visualizations of $\runs(\prog_{\ref{example-otp}})$ (L) and
    $\runs(\prog_{\ref{example-lambda-obliv}(a)})$ (R).}{fig-basic-distributions}
{\footnotesize
  $$
  \begin{array}{cc}
    \begin{array}{cccccc}
      \verb+s[1,0]+ & \verb+f[1,0]+ & \verb+v[2,0]+  & \verb+v[2,1]+ & \verb+v[2,2]+ & \verb+v[0,0]+\\
      \hline
      0 & 0 & 0 & 0 & 0 & 0 \\ 
      0 & 1 & 1 & 1 & 0 & 1 \\ 
      1 & 0 & 0 & 1 & 1 & 1 \\ 
      1 & 1 & 1 & 0 & 1 & 0
    \end{array}
    & 
    \begin{array}{ccccc}
      \verb+s[1,s]+ & \verb+f[1,sx]+ & \verb+f[1,sy]+ & \verb+v[0,0]+ & \verb+v[0,1]+ \\
      \hline
      0 & 0 & 0 & 0 & 0 \\ 
      0 & 0 & 1 & 1 & 0 \\ 
      0 & 1 & 0 & 0 & 1 \\ 
      0 & 1 & 1 & 1 & 1 \\
      1 & 0 & 0 & 0 & 0 \\ 
      1 & 0 & 1 & 0 & 0 \\ 
      1 & 1 & 0 & 1 & 1 \\ 
      1 & 1 & 1 & 1 & 1  
    \end{array}
  \end{array}
  $$
}
\end{fpfig}

%The basic distribution of a program $\prog$ is based on the
%final memories resulting from any run of $\prog$, denoted
%$\runs(\prog)$.
%\begin{definition}
%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
%  $$
%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
%  $$
%\end{definition}

As we've observed, $\runs(\prog)$ can be represented in a truth table
form, as illustrated in Figure \ref{fig-basic-distributions}.

\begin{fpfig}[t]{Filtering memories that satisfy a boolean expression.}{fig-solve}
{\small
\begin{eqnarray*}
\solve{\stores}{\etrue} &=& \stores\\
\solve{\stores}{\efalse} &=& \varnothing\\
\solve{\stores}{\flip{\cid}{w}} &=& \{ \store \in \stores \mid \store(\flip{\cid}{w}) \} \\
\solve{\stores}{\secret{\cid}{w}} &=& \{ \store \in \stores \mid \store(\secret{\cid}{w}) \} \\
\solve{\stores}{\view{\cid}{w}} &=& \{ \store \in \stores \mid \store(\view{\cid}{w}) \} \\
\solve{\stores}{\oracle{w}} &=& \{ \store \in \stores \mid \store(\oracle{w}) \} \\
\solve{\stores}{(\enot\ \be)} &=& \stores - (\solve{\stores}{\be})\\
\solve{\stores}{(\be_1\ \eand\ \be_2)} &=& (\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2}) \\
\solve{\stores}{(\be_1\ \eor\ \be_2)} &=& (\solve{\stores}{\be_1}) \cup (\solve{\stores}{\be_2}) \\
\solve{\stores}{(\be_1\ \exor\ \be_2)} &=&
 ((\solve{\stores}{\be_1}) \cap (\stores - \solve{\stores}{\be_2})) \cup\\
 && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \\
\solve{\stores}{\select{\be_1}{\be_2}{\be_3}} &=&
 ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup \\
 && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3})) \\
\solve{\stores}{\OT{\be_1}{\be_2}{\be_3}} &=&
 ((\solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_2})) \cup\\
 && ((\stores - \solve{\stores}{\be_1}) \cap (\solve{\stores}{\be_3}))
\end{eqnarray*}
}
\end{fpfig}

Here we describe a straightforward technique for obtaining program
distributions $\progd(\prog)$ through direct computation of
$\runs(\prog)$. Although we could obtain $\runs(\prog)$ simply by
executing $\prog$ given all random tapes and input secrets, the method
we describe here is empirically more efficient (though still
exponential in the length of the random tape), and also supports the
translation to Datalog described in Section
\ref{section-bruteforce-datalog}. An implementation of these
techniques along with Examples from Sections
\ref{section-minicat-examples} and \ref{section-metalang-ygc} is
available online\footnote{URL redacted for double-blind review.}.

In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
which filters $\stores$ based on satisfaction of $\be$. This algorithm
is fundamental to our brute force methods. It's correctness is characterized
as follows. 
\begin{lemma}
  \label{lemma-solves}
  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
  for some $\cid$.
\end{lemma}
\begin{proof}
  Since we assume safety of programs we can assume that all variables in $\be$ have the
  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
  on $\be$. 
\end{proof}

In order to automatically obtain $\runs(\prog)$, we can do a left
folding of $\solve$ across $\prog$, inductively extending memories
with valid view assignments in the order of their definition.
We can then use this to compute $\progd(\prog)$. We denote
these computations as $\cruns$ and $\cprogd$ respectively.
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\begin{proof}
  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
\end{proof}
\begin{lemma}
  \label{lemma-cprogd}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
  $\cprogd(\prog)$ as the distribution $\pmf$
  where for all $\store \in \mems(S\cup F \cup V)$:
  $$
  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then:
  $$
  \cprogd(\prog) = \progd(\prog)
  $$
\end{lemma}
\begin{proof}
  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
\end{proof}

\subsection{Protocol Certification}

### 2023-07-11 popl24/examples.tex
checkpoint

\section{Probability Mass Functions and Program Distributions}
\label{section-pmf}

Given a program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
\mems(\iov(\prog))$, we are fundamentally concerned with the joint
probability of the assignments in $\store_V$ conditioned by the
assignments $\store_S$ in any given run, since probabilistic
dependence between honest secrets and corrupt views is our metric of
information leakage.

\subsection{Basic Definitions}

We begin by defining joint probability mass functions in the standard
manner, though following \cite{barthe2019probabilistic} we use
memories to denote mappings of variables to values (i.e., outcomes),
so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
program containing these variables.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for some $X$ to values in $[0,1]$, such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ the \emph{marginal distribution} of variables $X$
  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
  $$
  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
  $$
\end{definition}

\begin{definition}
  Given $\margd{\pmf}{X}$, its \emph{conditional distribution given
  $\stores$} with $\dom(\stores) = Y$ with $Y \subseteq X$, denoted
  $\condd{\pmf}{X}{\stores}$, is a distribution with domain $X$ where for all
  $\store \in \mems(X)$:
  $$
  \condd{\pmf}{X}{\stores}(\store) =
  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
  $$
  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
\end{definition}
To recover familiar notation we allow the syntactic
sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
%\begin{eqnarray*}
%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
%\end{eqnarray*}

\subsection{Program Distributions}
Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Since $\fedcat$ is deterministic the
results (view assignments) of any run are determined by the input
secrets together with the random tape. And since we constrain programs
to not overwrite views, we are assured that \emph{final} memories
contain both a complete record of all initial secrets as well as views
resulting from communicated information.

We consider random tapes to be selected from a uniform distribution.
In our threat model, we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ and $\flips(\prog) = F$ we will consider all $\store
\in \mems(S \cup V \cup F)$ such that
$
\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this, we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pmf(\store) =  1 / |\stores| \ \text{if}\ \store \in \stores, \text{otherwise}\ 0
  %\begin{cases}
  %  1 / |\stores| & \text{if}\ \store \in \stores\\
  %  0 & \text{otherwise}
  %\end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution, i.e., 
  $
  \progd(\prog) \defeq  \margd{\pmf}{S\cup V}
  $.
  In some cases, we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progd(\prog, \be)$ to denote the program
  distribution $\progd(\prog;\eassign{\itv}{\be})$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

\subsection{Examples}
\label{section-pmf-examples}

>>>>

\section{Supplemental Examples}

\subsection{$\metaprot$ and $\minicat$ Protocol Examples}

\begin{example}[Additive Secret Sharing]
    \label{example-he}
Additive secret sharing is an MPC protocol where $k$ parties split up
their secrets into \emph{shares}, with the property that all $k$
shares are required to reconstruct the secret and any fewer reveals
nothing. Additionally, this sharing enjoys additive homomorphism. In
the binary field it is well-known that we can extend our use
of $\ttt{xor}$ and flips to generate an arbitrary number of shares
with additive homomorphism.

For example, to generate shares for three-party additive sharing,
clients 1,2, and 3 can break up and distribute their shares to
each other as follows:
{\small
\begin{verbatimtab}
     v[2,"s1"] := flip[1,"1"] xor flip[1,"s1"] xor s[1,"1"];
     v[3,"s1"] := flip[1,"1"];
     v[1,"s2"] := flip[2,"1"] xor flip[2,"s2"] xor s[2,"1"];
     v[3,"s2"] := flip[2,"1"];
     v[1,"s3"] := flip[3,"1"] xor flip[3,"s3"] xor s[3,"1"];
     v[2,"s3"] := flip[3,"1"] \end{verbatimtab}
}
Then, assuming that party 0 is a ``public'' client, each party sums
its own shares, and then the sum of the sum of shares is revealed
in the output view $\vx{0}{output}$. By additive homomorphism of
$\ttt{xor}$, this output is the sum of the 3 secrets.
{\small
\begin{verbatimtab}
     v[0,"ss1"] := v[1,"s2"] xor v[1,"s3"] xor flip[1,"s1"];
     v[0,"ss2"] := v[2,"s1"] xor v[2,"s3"] xor flip[2,"s2"];
     v[0,"ss3"] := v[3,"s1"] xor v[3,"s2"] xor flip[3,"s3"];
     v[0,"output"] := v[0,"ss1"] xor v[0,"ss2"] xor v[0,"ss3"] \end{verbatimtab}
}
We note that, as for MPC protocols generally, the notion of
``secrecy'' here is subtle and not the same as for encryption as in
Example \ref{example-otp}, particularly when the threat model allows
protocol participants to be corrupt. For example, suppose that $1 \in C$, and
$\sx{1}{1}$ is $1$, and after running the protocol suppose that
$\vx{0}{output}$ is 0. Then the adversary knows that 2 and 3's secrets
are either 1 and 0, or 0 and 1, with 50/50 probability, and definitely
neither 0 and 0, nor 1 and 1. 
\end{example}

Buliding on Example \ref{example-he}, we can define a function
$\ttt{share3}$ that abstracts the process of splitting a given
client's secret into 3 separate shares.
\begin{example} \label{example-share3} The function $\ttt{share3}$ 
  splits a client's secret into 3 shares returned as a record
  with fields $\ttt{s1-3}$:
  {\small
  \begin{verbatimtab}
      share3(client, secretid) {
        let s1 = flip[client, "share1"] in
        let s2 = flip[client, "share2"] in
        let s3 = (s1 xor s2) xor s[client, "s:" || secretid] in
        {s1 = s1;s2 = s2;s3 = s3}
      } \end{verbatimtab}
  }
  Here is a $\metaprot$ program that uses this function definition:
  {\small
    \begin{verbatimtab}
      let shares = share3(1, "mysecret") in
      v[2,s1] := shares.s2;
      v[3,s1] := shares.s3 \end{verbatimtab}
  }
  which generates the following $\minifed$ program, as we formalize in Example \ref{example-share3-eval}
  below:
  {\small
  \begin{verbatimtab}
      v[2,s1] := flip[1, "share2"];
      v[3,s1] := flip[1, "share1"] xor flip[1, "share2"] xor s[1, "s:mysecret"] \end{verbatimtab}
  }
\end{example}

\begin{example}
  \label{example-share3-eval}
  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program and let 
  $\prog_{\ref{example-share3}}$ be the  $\minifed$ program defined
  in Example \ref{example-share3}. We refer to the latter as ``accumulated''
  by evaluation of the former in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
  \redxs \config{\prog_{\ref{example-share3}}}{\ttt{()}}$.
\end{example}

\begin{example}
  Given $\ttt{share3}$ as defined in Example \ref{example-share3} and
  annotation:
  $$\tas(\ttt{share3}) =  \ttt{cid(client)}\ *\ \ttt{string(sid)}$$
  the type of $\ttt{share3}$ is $\tas(\ttt{share3}) \rightarrow \tau,\varnothing$
  where $\tau$ is:
  $$
  \ttt{\{ s1 : bool[client]; s2 : bool[client]; s3 : bool[client] \}}
  $$
\end{example}

\subsection{Program Distribution Examples}
\label{section-pmf-examples}

### 2023-07-11 popl24/metalang.tex
checkpoint

\begin{fpfig}[t]{Syntax of $\metaprot$.}{fig-metaprot-syntax}
  {\small
    $$
    \begin{array}{rcl@{\hspace{8mm}}r}
      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
      %x &\in& \mathrm{EVar}\\
      %f &\in& \mathrm{FName}\\[2mm]
      e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
      & & \select{e}{e}{e} \mid 
      \send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
      & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
      \mid e.\flab \mid e\concat e \mid (e) \\[2mm]
      v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
      \mid \ttt{()} & \textit{values}\\[2mm]
           {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
    \end{array}
    $$
  }
\end{fpfig}

@@

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot-syntax}.  It includes a syntax of function
definitions and records, and values include client ids, identifier
strings, and boolean expressions. Expression forms allow dynamic
construction of boolean expression forms and view assignments. When
$\metaprot$ programs construct an $\fedprot$ assignment, a side effect
occurs whereby the assignment is added to the end of the $\fedprot$
program accumulated during evaluation.

@@

When we consider the examples of GMW and YGC in detail below, our
focus will be on developing a codebase that can be used to define
arbitrary circuits, i.e., complete and concrete protocols. Since
strings and identifiers can be constructed manually, and expressions
can occur inside assignments and boolean expression forms, function
definitions can generalize over $\fedprot$-level patterns to obtain
composable program units. \cnote{Cur the following example to save
  space?.}  As a simple example, consider 3 party secret sharing as
illustrated in Example \ref{example-he}. We can define a function
$\ttt{share3}$ that abstracts the process of splitting a given
client's secret into 3 separate shares.
\begin{example} \label{example-share3} The function $\ttt{share3}$ 
  splits a client's secret into 3 shares returned as a record
  with fields $\ttt{s1-3}$:
  {\small
  \begin{verbatimtab}
      share3(client, secretid) {
        let s1 = flip[client, "share1"] in
        let s2 = flip[client, "share2"] in
        let s3 = (s1 xor s2) xor s[client, "s:" || secretid] in
        {s1 = s1;s2 = s2;s3 = s3}
      } \end{verbatimtab}
  }
  Here is a $\metaprot$ program that uses this function definition:
  {\small
    \begin{verbatimtab}
      let shares = share3(1, "mysecret") in
      v[2,s1] := shares.s2;
      v[3,s1] := shares.s3 \end{verbatimtab}
  }
  which generates the following $\minifed$ program, as we formalize in Example \ref{example-share3-eval}
  below:
  {\small
  \begin{verbatimtab}
      v[2,s1] := flip[1, "share2"];
      v[3,s1] := flip[1, "share1"] xor flip[1, "share2"] xor s[1, "s:mysecret"] \end{verbatimtab}
  }
\end{example}

@@

\begin{fpfig}[t]{Evaluation contexts and operational semantics of $\metaprot$.}{fig-metaprot-semantics}
  {\small
    $$
    \begin{array}{rcl@{\hspace{3mm}}r}
      E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid  \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
      & & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E} \mid \OT{E}{e}{e} \\[1mm]
      & & \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid \select{E}{e}{e} \mid \select{v}{E}{e} \mid \\[1mm]
      & & \select{v}{v}{E} \mid \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
      & & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\concat e \mid v \concat E
    \end{array}
    $$
    \medskip
    $$
    \begin{array}{rcl@{\hspace{10mm}}r}
      \config{\prog}{\elet{x}{v}{e}} &\redx& \config{\prog}{e[v/x]}\\
      \config{\prog}{f(v_1,...,v_n)} &\redx&
      \config{\prog}{e[v_1/x_1,\ldots,v_n/x_n]} & 
      \codebase(f) = x_1,\ldots,x_n,\ e\\
      \config{\prog}{\{\ldots; \flab = v; \ldots\}.\flab} &\redx&
      \config{\prog}{v}\\
      \config{\prog}{w_1\concat w_2} &\redx& \config{\prog}{w_1w_2}\\
      \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
      \config{\prog}{\instr} &\redx& \config{\prog;\instr}{()}\\
      \config{\prog}{E[e]} &\redx& \config{\prog'}{E[e']} & \text{if}\ \config{\prog}{e} \redx \config{\prog'}{e'} 
    \end{array}
    $$
  }
\end{fpfig}

We define a small-step evaluation aka reduction relation $\redx$ in
Figure \ref{fig-metaprot-semantics}.  We write $\redxs$ to denote the
reflexive, transitive closure of $\redx$. Reduction is defined on
\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
where $\prog$ is the $\minifed$ program accumulated during evaluation.
In this definition we write $e[v/x]$ to denote the substitution of $v$
for free occurences of $x$ in $e$. The rules are mostly standard,
except when a concrete $\minifed$ assignment is encountered it is added
to the end of $\prog$.

The rules rely on a definition of \emph{evaluation contexts} $E$
allowing computation within a larger program context, where $E[e]$
denotes an expression with $e$ in the hole $[]$ of $E$. Evaluation
contexts include boolean expression forms, allowing generalization
and instantiation of compositional program elements.
\begin{example}
  \label{example-share3-eval}
  Let $\codebase,e_{\ref{example-share3}}$ be the $\metaprot$ program and let 
  $\prog_{\ref{example-share3}}$ be the  $\minifed$ program defined
  in Example \ref{example-share3}. We refer to the latter as ``accumulated''
  by evaluation of the former in the sense that $\config{\varnothing}{e_{\ref{example-share3}}}
  \redxs \config{\prog_{\ref{example-share3}}}{\ttt{()}}$.
\end{example}

@@

The type syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot-tsyntax}. It includes a weak form of dependency:
string types $\stringty{e}$ and client types $\cidty{e}$
are parameterized by expressions $e$ that precisely reflect the
type of the value. Boolean expression forms have the type
$\bet{e}$ indexed by expressions $e$ indicating the client
id type of the expression. The dependency is weak in the
sense that expressions in types are a strict subset of
expression forms- for any $\stringty{e}$ the expression $e$
is either a variable, a string, or a concatenation form,
and for any $\cidty{e}$ or $\bet{e}$ the expression $e$
is either a variable or a client id $\cid$.

@@

The $\TirName{FnT}$ and $\TirName{Appt}$ rules apply to function
definition and application respectively, and rely on the
definition of function input type annotations $\tas$ and
type term substitutions $\sigma$. 
\begin{definition}
  A \emph{function input type annotation} $\tas$ is a mapping from
  function names $f$ to type products $\tau_1 * \cdots * \tau_n$.
  A \emph{type term substitution} $\sigma$ is a mapping from
  $\minifed$ variables $x$ to values, where $\sigma(\tau)$ denotes
  the replacement of occurences of $x$ in $\tau$ with $\sigma(x)$. 
\end{definition}
We assume that input type annotations $\tas$ are provided by the
programmer for all function definitions. This guarantees that
$\metaprot$ type checking is straightforward and efficient.
Function types are of the form
$
(\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst)
$,
where $\viewst$ denotes the effect of the function on the residual
program.  The function type can be understood as a dependent $\Pi$
type, with every term variable bound. When applied, these variables
are instantiated with a type term substitution $\sigma$. In our
implementation, we adapt \emph{synthesis} as defined in Dependent ML \cite{10.1145/292540.292560}
to obtain this $\sigma$- essentially this is a match on the syntactic
structure of types and expressions. 
\begin{example}
  Given $\ttt{share3}$ as defined in Example \ref{example-share3} and
  annotation:
  $$\tas(\ttt{share3}) =  \ttt{cid(client)}\ *\ \ttt{string(sid)}$$
  the type of $\ttt{share3}$ is $\tas(\ttt{share3}) \rightarrow \tau,\varnothing$
  where $\tau$ is:
  $$
  \ttt{\{ s1 : bool[client]; s2 : bool[client]; s3 : bool[client] \}}
  $$
\end{example}

>>>>

@@

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot}.  It includes a syntax of function
definitions and records, and values include client ids, identifier
strings, and boolean expressions. Expression forms allow dynamic
construction of boolean expression forms and view assignments. When
$\metaprot$ programs construct an $\fedprot$ assignment, a side effect
occurs whereby the assignment is added to the end of the $\fedprot$
program accumulated during evaluation.

@@

When we consider the examples of GMW and YGC in detail below, our
focus will be on developing a codebase that can be used to define
arbitrary circuits, i.e., complete and concrete protocols. Since
strings and identifiers can be constructed manually, and expressions
can occur inside assignments and boolean expression forms, function
definitions can generalize over $\fedprot$-level patterns to obtain
composable program units.

@@

\begin{fpfig}[t]{$\metaprot$ syntax (T), evaluation contexts (M), and operational semantics (B).}{fig-metaprot}
    {\small
    $$
    \begin{array}{rcl@{\hspace{8mm}}r}
      \multicolumn{3}{l}{\flab \in \mathrm{Field},\   x \in \mathrm{EVar}, \  f \in \mathrm{FName}}\\[2mm]
      %x &\in& \mathrm{EVar}\\
      %f &\in& \mathrm{FName}\\[2mm]
      e &::=& b \mid \flip{e}{e} \mid \secret{e}{e} \mid \view{e}{e} \mid \oracle{e} \mid \enot\ e \mid e\ \eand\ e \mid e\ \exor\ e \mid & \textit{expressions}\\[0mm]
      & & \select{e}{e}{e} \mid 
      \send{\view{e}{e}}{e} \mid \send{\view{e}{e}}{\OT{e}{e}{e}} \mid e;e \mid \\[0mm]
      & & x \mid \elet{x}{e}{e} \mid f(e,\ldots,e) \mid \{ \flab = e; \ldots; \flab = e \}
      \mid e.\flab \mid e\concat e \mid (e) \\[2mm]
      v &::=& w \mid \cid \mid \be \mid \{ \flab = v;\ldots;\flab = v \} 
      \mid \ttt{()} & \textit{values}\\[2mm]
           {fn} &::=& f(x,\ldots,x) \{ e \} & \textit{functions}
    \end{array}
    $$
    }

    \rule{130mm}{0.5pt}

    {\small
    $$
    \begin{array}{rcl@{\hspace{3mm}}r}
      E &::=& [\,] \mid \enot\ E \mid E\ \bop\ e \mid v\ \bop\ E \mid  \flip{E}{e} \mid \secret{E}{e} \mid \view{E}{e} \mid \oracle{E} \mid  \\[1mm]
      & & \flip{\cid}{E} \mid \secret{\cid}{E} \mid \view{\cid}{E} \mid \send{E}{e} \mid \send{\view{\cid}{w}}{E} \mid \OT{E}{e}{e} \\[1mm]
      & & \mid \OT{v}{E}{e} \mid \OT{v}{v}{E} \mid \select{E}{e}{e} \mid \select{v}{E}{e} \mid \\[1mm]
      & & \select{v}{v}{E} \mid \elet{x}{E}{e} \mid f(v,\ldots,v,E,e,\ldots,e) \mid \\[1mm]
      & & \{ \flab = v;\ldots;\flab = v;\flab = E;\flab = e;\ldots;\flab = e \} \mid E.\flab \mid E\concat e \mid v \concat E
    \end{array}
    $$
    \vspace{.5mm}

    \rule{130mm}{0.5pt}

    $$
    \begin{array}{rcl@{\hspace{10mm}}r}
      \config{\prog}{\elet{x}{v}{e}} &\redx& \config{\prog}{e[v/x]}\\
      \config{\prog}{f(v_1,...,v_n)} &\redx&
      \config{\prog}{e[v_1/x_1,\ldots,v_n/x_n]} & 
      \codebase(f) = x_1,\ldots,x_n,\ e\\
      \config{\prog}{\{\ldots; \flab = v; \ldots\}.\flab} &\redx&
      \config{\prog}{v}\\
      \config{\prog}{w_1\concat w_2} &\redx& \config{\prog}{w_1w_2}\\
      \config{\prog}{v;e} &\redx& \config{\prog}{e}\\
      \config{\prog}{\instr} &\redx& \config{\prog;\instr}{()}\\
      \config{\prog}{E[e]} &\redx& \config{\prog'}{E[e']} & \text{if}\ \config{\prog}{e} \redx \config{\prog'}{e'} 
    \end{array}
    $$
  }
\end{fpfig}

We define a small-step evaluation aka reduction relation $\redx$ in
Figure \ref{fig-metaprot}.  We write $\redxs$ to denote the
reflexive, transitive closure of $\redx$. Reduction is defined on
\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
where $\prog$ is the $\minifed$ program accumulated during evaluation.
In this definition we write $e[v/x]$ to denote the substitution of $v$
for free occurences of $x$ in $e$. The rules are mostly standard,
except when a concrete $\minifed$ assignment is encountered it is added
to the end of $\prog$.

The rules rely on a definition of \emph{evaluation contexts} $E$
allowing computation within a larger program context, where $E[e]$
denotes an expression with $e$ in the hole $[]$ of $E$. Evaluation
contexts include boolean expression forms, allowing generalization
and instantiation of compositional program elements.

@@

The type syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot-tsyntax}. It includes a weak form of dependency:
string types $\stringty{e}$ and client types $\cidty{e}$
are parameterized by expressions $e$ that precisely reflect the
type of the value. Boolean expression forms have the type
$\bet{e}$ indexed by expressions $e$ indicating the client
id type of the expression. The dependency is weak in the
sense that expressions in types are a strict subset of
expression forms- for any $\stringty{e}$ the expression $e$
is either a variable, a string, or a concatenation form,
and for any $\cidty{e}$ or $\bet{e}$ the expression $e$
is either a variable or a client id $\cid$. We define
the equivalence $w_1 \concat w_2 \equiv w_1w_2$ in typings.

@@

The $\TirName{FnT}$ and $\TirName{Appt}$ rules apply to function
definition and application respectively, and rely on the
definition of function input type annotations $\tas$ and
type term substitutions $\sigma$. 
\begin{definition}
  A \emph{function input type annotation} $\tas$ is a mapping from
  function names $f$ to type products $\tau_1 * \cdots * \tau_n$.
  A \emph{type term substitution} $\sigma$ is a mapping from
  $\minifed$ variables $x$ to values, where $\sigma(\tau)$ denotes
  the replacement of occurences of $x$ in $\tau$ with $\sigma(x)$. 
\end{definition}
We assume that input type annotations $\tas$ are provided by the
programmer for all function definitions. This guarantees that
$\metaprot$ type checking is straightforward and efficient.
Function types are of the form
$
(\tau_1 * \cdots * \tau_n \rightarrow \tau,\viewst)
$,
where $\viewst$ denotes the effect of the function on the residual
program.  The function type can be understood as a dependent $\Pi$
type, with every term variable bound. When applied, these variables
are instantiated with a type term substitution $\sigma$. In our
implementation, we adapt \emph{synthesis} as defined in Dependent ML \cite{10.1145/292540.292560}
to obtain this $\sigma$- essentially this is a match on the syntactic
structure of types and expressions.
